---
title: "The Wrath of God?"
author: "Jack Carter"
date: "6/4/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(tidyverse)
library(gutenbergr)
library(stringr)
library(ggplot2)
library(ggthemes)
library(tibble)
library(jsonlite)
library(tidytext)

# my personal plot theme for data visualizations. 
my_theme <- theme_economist_white(gray_bg = FALSE) +
  theme(plot.title = element_text(hjust = 0.5,
                                  vjust = 10,
                                  size = 10,
                                  color = "#474747"),
        plot.margin = unit(c(1.5, 1, 1.5, 1), "cm"),
        axis.text = element_text(size = 9,
                                 color = "gray30"),
        axis.text.x=element_text(vjust = -2.5),
        axis.title.x = element_text(size = 9,
                                    color = "gray30",
                                    vjust = -10),
        axis.title.y = element_text(size = 9,
                                    color = "gray30",
                                    vjust = 10),
        legend.direction = "vertical", 
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 11,
                                   color = "gray20"),
        legend.margin=margin(1, -15, 1, 0),
        legend.spacing.x = unit(0.25, "cm"),
        legend.key.size = unit(1, "cm"), 
        legend.key.height = unit(0.75, "cm"),
        strip.text = element_text(hjust = 0.5,
                                  vjust = 1,
                                  size = 10,
                                  color = "#474747"),
        panel.spacing = unit(2, "lines"))

# opens the cleaned books data frame.
df <- read_csv("final_df.csv") %>%
  mutate(book = str_replace_all(book, c("bible" = "the bible", 
                                        "quran" = "the quran")))

# breaks the text down into words and filters 
# out words of less than 2 characters. 
words <- df %>%
  unnest_tokens(word, text) %>%
  filter(nchar(word)>2)

# filters out stopwords.
words <- words %>%
  filter(!word %in% stop_words$word)

# joins the words with their sentiment values. 
sentiment <- inner_join(words, get_sentiments("nrc"))

# gets all the lines for a specific term. 
get_lines <- function(term) {
  lines <- words %>%
    summarize(words[words$word == term,"line_no"]) %>%
    mutate(label = term)
  return(lines)
}

# gets all the lines for a list of terms. 
get_multiple_terms <- function(terms) {
  lines <- list()
  for(i in 1:length(terms))
    lines[[i]] <- get_lines(terms[i])
  df <- tibble(rbind_pages(lines))
  return(df)
}

# creates a barplot of the anger for a list of terms.
get_anger <- function(data, terms, book_Type=NULL) {
  plot <- data %>% 
    filter(book_type == book_Type) %>%
    left_join(terms, by = "line_no") %>%
    filter(line_no %in% terms$line_no) %>%
    mutate(label = str_to_title(str_replace_all(label, "_", " ")),
           book = str_to_title(book),
           book = str_replace_all(book, c("The Bible" = "Bible", 
                                          "The Quran" = "Quran"))) %>%
    group_by(book, label) %>%
    mutate(total = length(sentiment)) %>%
    summarize(Anger = ((sum(sentiment=="anger")/total)*100)) %>%
    unique() %>%
    ggplot(aes(x=book,y=Anger,fill=book)) +
    ggtitle(str_to_title(str_replace_all(deparse(substitute(terms)),"_"," "))) +
    geom_bar(stat = "identity", position = "dodge") +
    facet_wrap(~label) +
    xlab("") +
    ylab("Anger (%)") +
    scale_fill_manual(values = c("#00B6EB", "#FB61D7")) +
    my_theme +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank())
  return(plot)
}

```

## **Summary**

The holy books, particularly the Quran, are often said to be full of hate. This project uses sentiment analysis to show how they are not necessarily full of hate, but rather very emotional. As with the priest in The Wrath of God, this means they may be prone to anger, but also contain a lot of positive emotions as well. 

&nbsp;

## **1) Anger**

1.1) The Quran is more angry than the Bible.  

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# visualizes anger. 
anger <- sentiment %>%
  filter(book_type == "biblical") %>%
  group_by(book) %>%
  summarize(Anger = (sum(sentiment=="anger")/length(sentiment)*100)) 

anger_df <- sentiment[,c("book", "book_type")] %>% 
  filter(book_type == "biblical") %>%
  left_join(anger,  by = "book") %>%
  distinct() %>%
  mutate(book_type = str_to_title(str_replace_all(book_type, "_", " ")),
         book = str_to_title(str_replace_all(book, "_", " ")))

anger_df %>%
  ggplot(aes(x=book,y=Anger,fill=book)) +
  ggtitle("Total Anger") +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("") +
  ylab("Anger (%)") +
  scale_fill_manual(values = c("#00B6EB", "#FB61D7")) +
  my_theme +
  theme(legend.position = "none",
        plot.margin = unit(c(2.5, 2.5, 2.5, 2.5), "cm"))
  

```

&nbsp;

1.2) The Quran is also more angry when speaking about the Israelite people. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

xenophobic_anger <- get_multiple_terms(c("jew","israel","hebrew",
                                        "jerusalem","jews","synagogue"))

get_anger(sentiment, xenophobic_anger, book_Type="biblical")

```

&nbsp;

1.3) The Quran is also more angry when speaking about sexual sins.  

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# sexual sins. 
sex <- get_multiple_terms(c("adultery","sodom","gomorrah",
                            "harlot","whore","fornication"))

get_anger(sentiment, sex, book_Type="biblical")

```

&nbsp;

## **2) Context**

2.1) The holy books are not that angry compared to other popular books. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# visualizes anger. 
anger <- sentiment %>%
  group_by(book) %>%
  summarize(Anger = (sum(sentiment=="anger")/length(sentiment)*100)) 

anger_df <- sentiment[,c("book", "book_type")] %>% 
  left_join(anger,  by = "book") %>%
  distinct() %>%
  mutate(book_type = str_to_title(str_replace_all(book_type, "_", " ")),
         book = str_to_title(str_replace_all(book, "_", " ")))

anger_df %>%
  ggplot(aes(x=book,y=Anger,fill=book)) +
  ggtitle("Total Anger") +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~book_type,
             scales = "free_x") +
  xlab("") +
  ylab("Anger (%)") +
  my_theme +
  theme(legend.position = "none") +
  scale_x_discrete(guide = guide_axis(n.dodge=3))

```

&nbsp;

2.2) Their apparent anger may stem from the fact that they are very emotional. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# visualizes how emotional the books are. 
get_hyperbole <- function(book) {
  emotions <- length(sentiment[sentiment$book==book,]$sentiment)
  total <- length(words[words$book==book,]$word)
  ratio <- (emotions/total)*100
  return(ratio)
}

books <- unique(sentiment$book)

hyperbole <- tibble(book = books,
                    hyperbole = unlist(lapply(books, get_hyperbole)))

hyperbole_df <- words[,c("book","book_type")] %>% 
  left_join(hyperbole, by = "book") %>%
  distinct() %>%
  mutate(book_type = str_to_title(str_replace_all(book_type, "_", " ")),
         book = str_to_title(str_replace_all(book, "_", " ")))

hyperbole_df %>%
  ggplot(aes(x=book,y=hyperbole,fill=book)) +
  ggtitle("Total Emotion") +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~book_type,
             scales = "free_x") +
  xlab("") +
  ylab("Emotional Words (%)") +
  my_theme +
  theme(legend.position = "none") +
  scale_x_discrete(guide = guide_axis(n.dodge=3))

```

&nbsp;

2.3 Their net sentiment is actually higher than most other popular books. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# visualizes net sentiment. 
net_senitment <- sentiment %>%
  group_by(book) %>%
  mutate(total = length(sentiment)) %>%
  summarize(Net=((sum(sentiment=="positive")/total)-((sum(sentiment=="negative"))/total))*100) %>%
  distinct()

net_senitment_df <- sentiment[,c("book", "book_type")] %>% 
  left_join(net_senitment,  by = "book") %>%
  distinct() %>%
  mutate(book_type = str_to_title(str_replace_all(book_type, "_", " ")),
         book = str_to_title(str_replace_all(book, "_", " ")))

net_senitment_df %>%
  ggplot(aes(x=book,y=Net,fill=book)) +
  ggtitle("Net Sentiment") +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~book_type,
             scales = "free_x") +
  xlab("") +
  ylab("Net Sentiment (%)") +
  my_theme +
  theme(legend.position = "none") +
  scale_x_discrete(guide = guide_axis(n.dodge=3))

```

&nbsp;

## **Disclaimer**

While the data above may suggest the Quran is more angry than the Bible, it is important to remember that it is also more emotional and actually has a higher net sentiment. Additionally, any conclusions made from sentiment analysis alone are limited, as it does not take context into account. 

&nbsp;

## **Method**

### **1) Download Books:**

The holy books and a selection of others are downloaded from the Project Gutenberg library.

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, eval=FALSE, message = FALSE, warning = FALSE}

# reads in the holy books from the gutenberg project library.
bible <- gutenberg_download(30) %>%
  mutate(gutenberg_id = "bible")

...

all_books <- bible %>%
  rbind(quran) %>%
  rbind(pride_and_prejudice) %>%
  rbind(communist_manifesto) %>%
  rbind(sherlock_holmes) %>%
  rbind(alice_in_wonderland) %>%
  rbind(the_great_gatsby)

```

&nbsp;

### **2) Clean Books:**

The holy books and a selection of others are cleaned by removing empty strings, punctuation and digits, as well as adding a column for biblical vs non-biblical and line number.  

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, eval=FALSE, message = FALSE, warning = FALSE}

# renames the columns. 
names(all_books) <- c("book","text")

# drops all the empty strings. 
all_books <- all_books[all_books$text !="",]

# cleans all digits and punctuation. 
all_books$text <- gsub('[[:punct:]]|[[:digit:]]','',all_books$text)

# adds a biblical/not biblical and line number column. 
final_df <- all_books %>%
  mutate(book_type = ifelse(book %in% c("bible", "quran"), "biblical", "non_biblical"),
         line_no = row_number())

```

&nbsp;

### **3) Get Sentiment:**

Each line is broken down into individual words (tokens), stop words (common words with little sentiment value) removed and the NRC sentiment library added via an inner join. 

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, eval=FALSE, message = FALSE, warning = FALSE}

# breaks the text down into words and filters 
# out words of less than 2 characters. 
words <- df %>%
  unnest_tokens(word, text) %>%
  filter(nchar(word)>2)

# filters out stopwords.
words <- words %>%
  filter(!word %in% stop_words$word)

# joins the words with their sentiment values. 
sentiment <- inner_join(words, get_sentiments("nrc"))

```

&nbsp;

### **4) Transformation:**

Various techniques are used to manipulate the data for the purpose of uncovering underlying trends. 

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, message = FALSE, warning = FALSE}

# extracts all words in line numbers matching a certain term. 
get_lines <- function(term) {
  lines <- words %>%
    summarize(words[words$word == term,"line_no"]) %>%
    mutate(label = term)
  return(lines)
}

get_multiple_terms <- function(terms) {
  lines <- list()
  for(i in 1:length(terms))
    lines[[i]] <- get_lines(terms[i])
  df <- tibble(rbind_pages(lines))
  return(df)
}

```

&nbsp;

### **5) Visualization:**

Finally, the data is visualized with barplots in ggplot2. 

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, message = FALSE, warning = FALSE}

# plots the anger associated with certain terms in the holy books. 
get_anger <- function(data, terms, book_Type=NULL) {
  plot <- data %>% 
    filter(book_type == book_Type) %>%
    left_join(terms, by = "line_no") %>%
    filter(line_no %in% terms$line_no) %>%
    mutate(label = str_to_title(str_replace_all(label, "_", " ")),
           book = str_to_title(book)) %>%
    group_by(book, label) %>%
    mutate(total = length(sentiment)) %>%
    summarize(Anger = ((sum(sentiment=="anger")/total)*100)) %>%
    unique() %>%
    ggplot(aes(x=book,y=Anger,fill=book)) +
    ggtitle(str_to_title(str_replace_all(deparse(substitute(terms)),"_"," "))) +
    geom_bar(stat = "identity", position = "dodge") +
    facet_wrap(~label) +
    xlab("") +
    ylab("Sentiment (%)") +
    my_theme +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  return(plot)
}

```

&nbsp;

## **Sources**

- Firestone (2020) https://www.degruyter.com/document/doi/10.1515/9783110671773-007/html

- Project Gutenberg (2022) https://www.gutenberg.org/

- Mohammad (2022) https://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm

&nbsp;
&nbsp;
&nbsp;
